Stability and accuracy of free surface time integration in viscous flows

### 2016-11-08 new_response_to_referee.tex

\begin{quotation}
Introduction. The earliest implementation of sticky air approach was published by Matsumoto, T., Tomoda, Y., 1983. Numerical-simulation of the initiation of subduction at the fracture-zone. J. Phys. Earth 31, 183–194.
Introduction. There is one recent paper about true free surface implementation with staggered grid by T Duretz, DA May, P Yamato  (2016) A free surface capturing discretization for the staggered grid finite difference scheme Geophysical Journal International 204 (3), 1518-1530 
\end{quotation}
We have added these pertinent references.

\begin{quotation}
Line 27. “A true free surface has mathematical elegance in that the boundary condition of the domain more closely matches the boundary conditions which one is trying to model, but it typically requires a deformable domain with frequent remeshing to avoid ill-conditioned cells.” One should mention here, however, possibilities of treating more complex realistic cases of common geodynamic simulations, in which oceanic free surface is covered by water whereas continental regions contact with the athmosphere. Density contrast along such free surface will be non-uniform, which would require some special treatments.
\end{quotation}
This is a good point, and we have added a note to this effect.
\begin{quotation}
Equations 1,3,4,5 – minus is missing in front of RHO*g
\end{quotation}
These are now fixed.

\begin{quotation}
The authors mentioned that they did Schmeling et al. benchmark (2008) but do not show it since it is difficult to interpret and quantify. I would still recommend showing one-two figures with the progress of model evolution and surface development for this model. It will simply show that the developed algorithm does not break when treating more complex plate tectonics-like models with large viscosity contrasts and complex evolution and advection of the free surface.
\end{quotation}

We investigated the subduction benchmark of Schmeling et al. 2008, but do not report
the results because the benchmark is not well designed to test our time-stepping scheme.
The chief difficulty is sharp corners in the shape of the initial slab. While the initial condition
is easy to implement, it introduces singularities into the problem. These singularities are
resolved indirectly by limiting the spatial resolution. Moreover the viscosity contrast in this
subduction benchmark is only a factor of 100. Relatively little is gained over the (smaller)
factor of 10 viscosity increase in the Rayleigh-Taylor benchmark.

### 2016-11-11 free-surface-paper.tex
Add Taras to thank yous.
add sentence to abstract

Finally, we discuss the implementation of a free surface in the open source, community based
mantle convection software \texttt{ASPECT}.
\end{abstract}

@@

\section*{Acknowledgements}
This work was supported through National Science Foundation grant EAR-1246670
and by the Computational Infrastructure in Geodynamics (CIG), through National Science Foundation grant EAR-0949446. 
TH is partially supported by NSF grant DMS-1522191.
Figures were created using the Python packages NumPy and Matplotlib.
Thanks to Wolfgang Bangerth, Brent Delbridge, and Cedric Thieulot for useful discussions.
Thanks to Fabio Crameri and Dave May for constructive reviews.

>>>>

Finally, we present numerical results to show the effectiveness of the new approach and discuss the free surface implementation in the open source, community based
mantle convection software \texttt{ASPECT}.
\end{abstract}

@@

\section*{Acknowledgements}
This work was supported through National Science Foundation grant EAR-1246670
and by the Computational Infrastructure in Geodynamics (CIG), through National Science Foundation grant EAR-0949446. 
TH is partially supported by NSF grant DMS-1522191.
Figures were created using the Python packages NumPy and Matplotlib.
Thanks to Wolfgang Bangerth, Brent Delbridge, and Cedric Thieulot for useful discussions.
Thanks to Fabio Crameri, Taras Gerya, and Dave May for constructive reviews.
